The effect of climate forcing on numerical simulations of the Cordilleran Ice Sheet at the Last Glacial Maximum

### 2013-10-31 sections/discuss.tex
New discussion first draft

@@

\subsection{Potential method improvements}

>>>>

The outcome of our numerical simulations shows a very large sensitivity to the choice of climate forcing data (Fig.~\ref{fig:ivolarea}-\ref{fig:cool05}).To understand the origin of these discrepancies, we compare the distributions of surface air temperature and precipitation rate among between datasets and test their effect on modelled ice sheet geometries. We then compare our findings to a reconstruction of the LGM Cordilleran Ice Sheet margin, and discuss some of the key assumptions of this study.

\subsection{Comparison of forcing climatologies}

The PDD model that we use is potentially highly sensitive to temperature and precipitation forcing. More particularly, winter (DJF) precipitation and summer (JJA) temperature respectively predominantly drive accumulation and melt at the surface of the modelled ice sheet. As illustrated (Fig.~\ref{fig:topo}-\ref{fig:prec}), significant differences between datasets exist in their distribution exist.

Surface air temperature from the WorldClim observational data is fairly well resolved by reanalysis data (Fig.~\ref{fig:tempheatmap}), apart of the NCAR reanalysis which shows a large spread around the WorldClim values. For ERA-Interim, NARR and CFSR reanalyses, it seems that most of the differences are due to unresolved local topographic detail (Fig.~\ref{fig:tempdiff}). NCAR shows a significant cold anomaly over the entire region. It is known from literature that NCAR produces too cold temperature at high latitudes. Finally NARR data seem to be a bit too warm.

Regarding precipitation data, more significant differences exist between datasets. All reanalysis data used seem to overestimate precipitation rates over most of the modelling domain (Fig.~\ref{fig:precheatmap}). The overestimation is the largest for CFSR data. Again, NCAR shows the largest discrepancy and this is probably due to the coarse model resolution. When looking at the spatial distribution of precipitation anomalies, it appears that precipitation is underestimated on most of the major reliefs, while it is overestimated on the lowlands and plateaux located beyond these reliefs (Fig.~\ref{fig:precdiff}). We think this is an orographic precipitation effect linked to horizontal GCM resolution (Fig.~\ref{fig:oroprecip}). Precipitation rate, however, is generally and largely overestimated in CFSR data, despite of high model resolution.

To distinguish the role of temperature and precipitation biases outlined above in forcing the ice sheet model, we run a series of additional experiments using artificial, “hybrid” atmospheres that consist of temperature data from WorldClim, and precipitation data from each of the four reanalysis, and the other way around. These eight simulations were run using a temperature offset value of 5\,K, and temperature lapse rates were computed using the reference topography from the appropriate data.

Although only temperature or precipitation changes are applied, this experiment again resulted in a largely different modelled ice sheet geometries (Fig.~\ref{fig:biatm}). Varying temperature forcing alone however, had limited effect, expect in the case of NCAR, due to its cold bias previously mentioned. For ERA-Interim, NARR and CFSR data, variations in precipitations dominate the ice sheet response, whereas for NCAR data, both temperature and precipitation are responsible for the anomalously large ice sheets produced (Fig.~\ref{fig:biatmbars}).

@@

\subsection{Sensitivity to englaciation duration}

Among the most decisive assumptions for the present study figures certainly the hypothesis that the Cordilleran Ice Sheet built up from ice free conditions to LGM conditions in 10Ka, which motivates our choice of simulation length. Although geomorphological data shows that this period could not have been much longer, they provide no lower bound for it. The growth of the Cordilleran Ice Sheet to full glacial condition could have lasted less than 10Ka.

In our simulations, using NARR data and temperatures more than 7 degrees below present, the modelled ice sheet reaches near-LGM ice cover in less than 10Ka (Fig.~\ref{fig:durationstack}). A shorter glaciation period lead to fewer ice cover in the eastern part of the modelling domain, where precipitation is limiting the rate of growth, but further glaciation along the Pacific margin to the south, where warm temperatures normally stop glacier advance. This shows that our choice of 10Ka may be somewhat too much. On the other hand our representation of climate history is voluntarily simplistic, and the effects of a more complex palaeo-climate history on the Cordilleran Ice Sheet will need to be studied further.

\subsection{Other potential method improvements}
